Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated

### 2022-06-29 MScThesisText.tex
version before presentation

Worldwide, millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work. These are the so-called ``left-behind children'' or ``stay-behind children.'' They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children (DeWaard et al., 2018), one-fifth of Bulgarian children (Popova, 2018), and one-fourth of Chinese children were living apart from their migrating parents (NBS et al., 2017).

@@

A challenge arises when we want to assess the causality between parent's migration and children's educational achievement. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, we can infer causality from a quasi-experimental design -- that is, when the cause can be on par with random assignment to participants in the analysis, conditional on identification assumptions (Cunningham, 2021).

In this thesis, my research question is: To what extent do parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyzed data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and applied propensity score matching for robustness check. The outcome was measured with children's cognitive test score and academic exam scores. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic capital is transferable to human capital, which is the stock of ``skills, talents, health, and expertise'' in people that enhance their productivity (Botev et al., 2019; Goldin, 2016). D. Yang (2008) noted that in the Philippes, a sharp depreciation of peso increased the amount of remittances from migrant laborers working abroad. This subsequently boosted children's school enrollment and educational expenditure. In China, Chang et al. (2019) interviewed migrant workers and their family members. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood.

@@

Researchers have studied whether the impact differ between mother absence and father absence in the context of labor migration. Y. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was ``negatively associated with children's social competence and academic performance,'' while father migration had no direct effects on the two outcomes (pp.~860-861).

@@

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers (NBS, 2022). They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The number of left-behind Chinese children reached 68.7 million in 2015 (NBS et al., 2017). The internal migration stemmed from the neoliberal economic regime.

The neoliberal economic regime sounds at odds with China, a so-called socialist country. Actually, in the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the world economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. The regional inequality needs social policy to amend, but China's welfare programs are discriminatory -- it segregates the population. Those who reap the greatest benefits are public servants, urban dwellers, and formally employed people. As rural people migrate to the city, they find themselves with limited access to social welfare in the destination. The municipal governments only provide welfare to local residents and expect non-local residents access public services elsewhere (Duckett, 2020; Zhang, 2018).

The institutionalized discrimination against migrant laborers find its roots in China's ``multilevel citizenship'' . As Vortherms (2021) demonstrated, citizenship is defined by ``membership rules and rights entitlements;'' they operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through the \emph{hukou}, or household registration system:

@@

In other words, the hukou system binds one's citizenship rights with one's place of registry. One can move out of town freely but cannot change hukou freely. When a rural resident moves to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the city {[}Gu (2021); wang2020{]}. Migrant workers had organized strikes and protests, only to meet with police crackdown (Chan \& Selden, 2017). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns.

@@

The control variables on children include their age, gender, ethnicity, self-rated health, hukou type, number of siblings, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the level of education of the best-educated parent, the parental expectation on the child's level of education, the perceived household financial conditions, the number of extracurricular books at home, and the internet assess at home.

@@

Table \ref{i-var} below summarizes the changes in migration status in the sample. All 4088 children were living together with parents in one household in baseline, but about fifteen percent saw at least one parent migrated during the two waves of survey. In these new migrant households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth was both-parents migration.

@@

The tables below provide the whole-sample summary statistics. In addition, the violin plots in the appendix visualize the distribution of outcome variables for the treatment group and the control group.

\begin{longtable}[]{@{}lrrrr@{}}
\toprule
& Mean & SD & Min & Max \\
\midrule
\endhead
Cognitive score difference & 0.27 & 0.80 & -2.88 & 4.03 \\
Cognitive score at wave 1 & 0.15 & 0.85 & -2.03 & 2.33 \\
Cognitive score at wave 2 & 0.42 & 0.78 & -3.14 & 2.06 \\
Academic exam score difference & -8.82 & 33.21 & -207.50 & 166.00 \\
Academic exam score at wave 1 & 213.11 & 49.46 & 19.33 & 297.50 \\
Academic exam score at wave 2 & 204.29 & 55.41 & 0.00 & 293.33 \\
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}llrr@{}}
\toprule
& & N & \% \\
\midrule
\endhead
migr & 0 & 3452 & 84.4 \\
& 1 & 636 & 15.6 \\
child.is.girl & 0 & 2002 & 49.0 \\
& 1 & 2086 & 51.0 \\
child.has.rural.hukou & 0 & 2221 & 54.3 \\
& 1 & 1867 & 45.7 \\
child.in.boarding.school & 0 & 2906 & 71.1 \\
& 1 & 1182 & 28.9 \\
child.is.ethnic.minority & 0 & 3800 & 93.0 \\
& 1 & 288 & 7.0 \\
child.had.skipped.grade & 0 & 4042 & 98.9 \\
& 1 & 46 & 1.1 \\
child.had.repeated.grade & 0 & 3700 & 90.5 \\
& 1 & 388 & 9.5 \\
child.went.to.preschool & 0 & 639 & 15.6 \\
& 1 & 3449 & 84.4 \\
parent.has.white.collar.job & 0 & 3238 & 79.2 \\
& 1 & 850 & 20.8 \\
home.has.internet & 0 & 1316 & 32.2 \\
& 1 & 2772 & 67.8 \\
child.health.self.rated & Very poor & 21 & 0.5 \\
& Not very good & 113 & 2.8 \\
& Moderate & 807 & 19.7 \\
& Good & 1383 & 33.8 \\
& Very good & 1764 & 43.2 \\
child.educational.aspiration & Junior high school & 140 & 3.4 \\
& Senior high school & 418 & 10.2 \\
& Associate college & 616 & 15.1 \\
& University & 1205 & 29.5 \\
& Postgraduate & 1709 & 41.8 \\
parent.educational.level & Primary school and below & 248 & 6.1 \\
& Junior high school & 1580 & 38.6 \\
& Senior high school & 1254 & 30.7 \\
& Associate college & 385 & 9.4 \\
& University & 543 & 13.3 \\
& Postgraduate & 78 & 1.9 \\
parent.expectation.on.child & Junior high school & 36 & 0.9 \\
& Senior high school & 257 & 6.3 \\
& Associate college & 426 & 10.4 \\
& University & 1574 & 38.5 \\
& Postgraduate & 1795 & 43.9 \\
home.extracurricular.book & Very few & 395 & 9.7 \\
& Not many & 414 & 10.1 \\
& Some & 1312 & 32.1 \\
& Quite a few & 1111 & 27.2 \\
& A great number & 856 & 20.9 \\
home.economic.resource & Very poor & 115 & 2.8 \\
& Somewhat poor & 598 & 14.6 \\
& Moderate & 3124 & 76.4 \\
& Somewhat rich & 239 & 5.8 \\
& Very rich & 12 & 0.3 \\
\bottomrule
\end{longtable}

@@

The unrestricted DID models support my hypothesis one -- newly left-behind adolescents are more likely to perform worse in cognitive and academic abilities than their non left-behind counterparts. If a household sees any parent migrated, the child's cognitive test score and total exam grade would decrease relative to that of children whose parents stay with them, holding everything else constant. Parental migration is linked to is -0.09 SD in cognitive test score and -0.09 points in academic exam grades. Both relationships are slight in strength.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1398}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2043}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2151}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2151}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2258}}@{}}
\caption{Parental migration's effect on children's cognitive abilities}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Absent
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Absent
\end{minipage} \\
\midrule
\endhead
Explanator & -0.09** & -0.04 & -0.14** & -0.10 \\
& (0.03) & (0.05) & (0.04) & (0.06) \\
W1 score & -0.62*** & -0.63*** & -0.62*** & -0.62*** \\
& (0.02) & (0.02) & (0.02) & (0.02) \\
Num.Obs. & 4088 & 3706 & 3703 & 3583 \\
R2 & 0.494 & 0.502 & 0.505 & 0.499 \\
R2 Adj. & 0.424 & 0.426 & 0.430 & 0.419 \\
R2 Within & 0.370 & 0.378 & 0.378 & 0.374 \\
FE: clsids & X & X & X & X \\
FE: w2clsids & X & X & X & X \\
\bottomrule
\end{longtable}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1398}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2043}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2151}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2151}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2258}}@{}}
\caption{Parental migration's effect on children's academic abilities}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Absent
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Absent
\end{minipage} \\
\midrule
\endhead
Explanator & -2.67* & -5.85* & -0.25 & -1.13 \\
& (1.26) & (2.36) & (1.66) & (2.28) \\
W1 score & 0.04 & 0.04† & 0.04† & 0.04 \\
& (0.02) & (0.02) & (0.02) & (0.02) \\
Num.Obs. & 4088 & 3706 & 3703 & 3583 \\
R2 & 0.562 & 0.577 & 0.566 & 0.577 \\
R2 Adj. & 0.501 & 0.513 & 0.500 & 0.510 \\
R2 Within & 0.005 & 0.009 & 0.004 & 0.004 \\
FE: clsids & X & X & X & X \\
FE: w2clsids & X & X & X & X \\
\bottomrule
\end{longtable}

@@

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1398}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2043}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2151}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2151}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2258}}@{}}
\caption{Parental migration's effect on children's cognitive abilities, estimated with matching}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Absent
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Absent
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Absent
\end{minipage} \\
\midrule
\endhead
Explanator & -0.09** & -0.05 & -0.13* & -0.05 \\
& (0.03) & (0.06) & (0.06) & (0.11) \\
W1 score & -0.61*** & -0.65*** & -0.60*** & -0.63*** \\
& (0.03) & (0.04) & (0.04) & (0.06) \\
Num.Obs. & 1958 & 913 & 897 & 475 \\
R2 & 0.528 & 0.627 & 0.638 & 0.646 \\
R2 Adj. & 0.375 & 0.313 & 0.320 & -0.312 \\
R2 Within & 0.363 & 0.407 & 0.372 & 0.351 \\
FE: clsids & X & X & X & X \\
FE: w2clsids & X & X & X & X \\
\bottomrule
\end{longtable}

@@

\begin{figure}
\centering
\includegraphics{MScThesisText_files/figure-latex/v-plot-ttl-1-1.pdf}
\caption{\label{fig:v-plot-ttl-1}Changes in standardized cognitive test score by household migration arrangement}
\end{figure}

@@

>>>>

Worldwide, millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work. These are the so-called ``left-behind children'' or ``stay-behind children.'' They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children (DeWaard et al., 2018), one-fifth of Bulgarian children (Popova, 2018), and one-fourth of Chinese children were living apart from their migrating parents (NBS et al., 2017). In a 2021 resolution, the Council of Europe's Parliamentary Assembly (2021) expressed concern about ``the scale of this phenomenon and the long-term damage it creates'' (p.~2) and declared that leaving ``millions of labour migrants' children without parental care is a mass violation of human rights'' (p.~1).

@@

A challenge arises in assessing the causality between parent's migration and children's educational achievement. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. The cause should be on par with random assignment to participants in the analysis, conditional on identification assumptions (Cunningham, 2021).

In this thesis, my research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyzed data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and applied propensity score matching for robustness check. The outcome was measured with children's cognitive test score and academic exam scores. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic capital is transferable to human capital, which is the stock of ``skills, talents, health, and expertise'' in people that enhance their productivity (Botev et al., 2019; Goldin, 2016). D. Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational expenditure. In China, Chang et al. (2019) interviewed migrant workers and their family members. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood.

@@

Researchers have studied whether the impact differ between mother absence and father absence induced by labor migration. Y. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was ``negatively associated with children's social competence and academic performance,'' while father migration had no direct effects on the two outcomes (pp.~860-861).

@@

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers (NBS, 2022). They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind (NBS et al., 2017). The internal migration stemmed from the neoliberal economic regime.

The neoliberal economic regime sounds at odds with China, a so-called socialist country. Actually, in the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. The regional inequality needs social policy to amend, but China's welfare program is discriminatory -- it segregates the population. Those who reap the greatest benefits are public servants, urban dwellers, and formally employed people. As rural people migrate to the city, they find themselves with limited access to social welfare in the destination. The municipal governments only provide welfare to local residents and expect non-local residents access public services elsewhere (Duckett, 2020; Zhang, 2018).

The institutionalized discrimination against migrant laborers find its roots in China's ``multilevel citizenship'' system. As Vortherms (2021) demonstrated, the constituents of citizenship -- ``membership rules and rights entitlements'' -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through the \emph{hukou}, or household registration system:

@@

In other words, the hukou system binds one's citizenship rights with one's place of registry. In theory, anyone can move out of town freely. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city (Gu, 2022). Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown (Chan, 2019). Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns.

@@

The control variables on children include their age, gender, ethnicity, self-rated health, hukou type, number of siblings, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the level of education of the best-educated parent, the parental expectation on the child's level of education, the perceived household financial conditions, the number of extracurricular books at home, and the internet access at home.

@@

Table 1 below summarizes the changes in migration status in the sample. All 4088 children were living together with parents in one household in baseline, but about fifteen percent saw at least one parent migrated during the two waves of survey. In these new migrant households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth was both-parents migration.

\renewcommand{\arraystretch}{0.75}

@@

Tables 2 below summarizes the dependent variables by the control group and the treatment group. In addition, the violin plots in the appendix visualize the distribution of outcome variables.

\begin{tabular}{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3452} & \multicolumn{2}{c}{Yes = 636} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & Mean & Sd & Mean & Sd & Test\\
\midrule
Cognitive test score difference & 0.28 & 0.8 & 0.19 & 0.8 & p=0.006**\\
\hspace{1em}Wave 1 score & 0.15 & 0.86 & 0.13 & 0.84 & p=0.471\\
\hspace{1em}Wave 2 score & 0.43 & 0.78 & 0.31 & 0.8 & p<0.001***\\
Academic exam score difference & -8.2 & 32.75 & -12.15 & 35.43 & p=0.006**\\
\hspace{1em}Wave 1 score & 213.39 & 49.31 & 211.58 & 50.3 & p=0.395\\
\addlinespace
\hspace{1em}Wave 2 score & 205.19 & 54.83 & 199.42 & 58.23 & p=0.016*\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}

\begin{tabular}{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3452} & \multicolumn{2}{c}{Yes = 636} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & N & Percent & N & Percent & Test\\
\midrule
child.is.girl &  &  &  &  & X2=11.942***\\
... 0 & 1650 & 48\% & 352 & 55\% & \\
... 1 & 1802 & 52\% & 284 & 45\% & \\
child.has.rural.hukou &  &  &  &  & X2=0.485\\
... 0 & 1884 & 55\% & 337 & 53\% & \\
\addlinespace
... 1 & 1568 & 45\% & 299 & 47\% & \\
child.in.boarding.school &  &  &  &  & X2=4.63*\\
... 0 & 2477 & 72\% & 429 & 67\% & \\
... 1 & 975 & 28\% & 207 & 33\% & \\
child.is.ethnic.minority &  &  &  &  & X2=0.388\\
\addlinespace
... 0 & 3213 & 93\% & 587 & 92\% & \\
... 1 & 239 & 7\% & 49 & 8\% & \\
child.had.skipped.grade &  &  &  &  & X2=0.02\\
... 0 & 3414 & 99\% & 628 & 99\% & \\
... 1 & 38 & 1\% & 8 & 1\% & \\
\addlinespace
child.had.repeated.grade &  &  &  &  & X2=10.621**\\
... 0 & 3147 & 91\% & 553 & 87\% & \\
... 1 & 305 & 9\% & 83 & 13\% & \\
child.went.to.preschool &  &  &  &  & X2=0.923\\
... 0 & 531 & 15\% & 108 & 17\% & \\
\addlinespace
... 1 & 2921 & 85\% & 528 & 83\% & \\
parent.has.white.collar.job &  &  &  &  & X2=0.001\\
... 0 & 2735 & 79\% & 503 & 79\% & \\
... 1 & 717 & 21\% & 133 & 21\% & \\
home.has.internet &  &  &  &  & X2=12.815***\\
\addlinespace
... 0 & 1072 & 31\% & 244 & 38\% & \\
... 1 & 2380 & 69\% & 392 & 62\% & \\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}

\begin{tabular}{llllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental Migration} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{No = 3452} & \multicolumn{2}{c}{Yes = 636} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & Mean & Sd & Mean & Sd & Test\\
\midrule
child.age & 12.45 & 0.63 & 12.5 & 0.72 & F=3.616\\
child.health.self.rated & 4.17 & 0.87 & 4.14 & 0.9 & F=0.622\\
child.educational.aspiration & 3.97 & 1.12 & 3.93 & 1.19 & F=0.617\\
parent.educational.level & 2.9 & 1.19 & 2.94 & 1.21 & F=0.367\\
parent.expectation.on.child & 4.19 & 0.91 & 4.15 & 0.94 & F=1.093\\
\addlinespace
home.extracurricular.book & 3.41 & 1.18 & 3.32 & 1.28 & F=3.09\\
home.economic.resource & 2.87 & 0.55 & 2.84 & 0.58 & F=1.37\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}* p < 0.05, ** p < 0.01, *** p < 0.001}\\
\end{tabular}

@@

The unrestricted DID models support my hypothesis one -- newly left-behind adolescents are more likely to perform worse in cognitive and academic abilities than their non left-behind counterparts. If a household sees any parent migrated, the child's cognitive test score and total exam grade would decrease relative to that of children whose parents stay with them, holding everything else constant. Parental migration is linked to -0.09 SD in cognitive test score and -0.09 points in academic exam grades. Both relationships are slight in strength.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1287}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2079}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2178}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2178}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2277}}@{}}
\caption{Parental migration's effect on children's cognitive test score}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Migrated
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Migrated
\end{minipage} \\
\midrule
\endhead
Migration & -0.09** & -0.04 & -0.14** & -0.10 \\
& (0.03) & (0.05) & (0.04) & (0.06) \\
W1 score & -0.62*** & -0.63*** & -0.62*** & -0.62*** \\
& (0.02) & (0.02) & (0.02) & (0.02) \\
Num.Obs. & 4088 & 3706 & 3703 & 3583 \\
R2 & 0.494 & 0.502 & 0.505 & 0.499 \\
R2 Adj. & 0.424 & 0.426 & 0.430 & 0.419 \\
R2 Within & 0.370 & 0.378 & 0.378 & 0.374 \\
FE: clsids & X & X & X & X \\
FE: w2clsids & X & X & X & X \\
\bottomrule
\end{longtable}

@@

\textbf{Note:}
\^{}\^{} Control variables not included

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1287}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2079}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2178}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2178}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2277}}@{}}
\caption{Parental migration's effect on children's academic exam score}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Migrated
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Migrated
\end{minipage} \\
\midrule
\endhead
Migration & -2.67* & -5.85* & -0.25 & -1.13 \\
& (1.26) & (2.36) & (1.66) & (2.28) \\
W1 score & 0.04 & 0.04† & 0.04† & 0.04 \\
& (0.02) & (0.02) & (0.02) & (0.02) \\
Num.Obs. & 4088 & 3706 & 3703 & 3583 \\
R2 & 0.562 & 0.577 & 0.566 & 0.577 \\
R2 Adj. & 0.501 & 0.513 & 0.500 & 0.510 \\
R2 Within & 0.005 & 0.009 & 0.004 & 0.004 \\
FE: clsids & X & X & X & X \\
FE: w2clsids & X & X & X & X \\
\bottomrule
\end{longtable}

@@

\textbf{Note:}
\^{}\^{} Control variables not included

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1287}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2079}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2178}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2178}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2277}}@{}}
\caption{Parental migration's effect on children's cognitive test scores, estimated with matching}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Migrated
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Any Parent Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Mother Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Only Father Migrated
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Both Parents Migrated
\end{minipage} \\
\midrule
\endhead
Migration & -0.09** & -0.05 & -0.13* & -0.05 \\
& (0.03) & (0.06) & (0.06) & (0.11) \\
W1 score & -0.61*** & -0.65*** & -0.60*** & -0.63*** \\
& (0.03) & (0.04) & (0.04) & (0.06) \\
Num.Obs. & 1958 & 913 & 897 & 475 \\
R2 & 0.528 & 0.627 & 0.638 & 0.646 \\
R2 Adj. & 0.375 & 0.313 & 0.320 & -0.312 \\
R2 Within & 0.363 & 0.407 & 0.372 & 0.351 \\
FE: clsids & X & X & X & X \\
FE: w2clsids & X & X & X & X \\
\bottomrule
\end{longtable}

@@

\begin{figure}
\centering
\includegraphics{MScThesisText_files/figure-latex/v-plot-ttl-1-1.pdf}
\caption{\label{fig:v-plot-ttl-1}Changes in academic exam score by household migration arrangement}
\end{figure}

@@

\leavevmode\vadjust pre{\hypertarget{ref-duxe9murger2015}{}}%
Démurger, S., \& Xu, H. (2015). Left-behind children and return migration in china. \emph{IZA Journal of Migration}, \emph{4}(1), 10. \url{https://doi.org/10.1186/s40176-015-0035-x}
